---
title: '1M03 Assignment #1'
author: "Roy Luo 400474680, Kenny Zhao"
date: "2023-01-30"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=TRUE)
set.seed(34)
library(dplyr)
```

## Question 1
```{r}
N <- 1000
m1 <- matrix(sample(1:10, N, replace=TRUE), ncol=5)
```
**a.**
```{r}
df1 <- as.data.frame(m1)
```
**b.**
```{r}
names(df1)
```
**c.**
```{r}
names(df1)<-paste0("col_", 1:ncol(df1))
names(df1)
```
**d.**
```{r}
lst1 <- list(m1,df1)
glimpse(lst1)
```
**e.**
```{r}
v1_ <- lst1[[1]][,3]
v2_ <- lst1[[2]][,3]
```
**f.**
```{r}
identical(v1_,v2_)
```
## Question 2
```{r}
N <- 50
x1 <- as.factor(sample(letters[7:12], N, replace=TRUE))
x2 <- as.factor(sample(7:12, N, replace = TRUE))
```
**a.**
```{r}
class(x1)
```
**b.**
```{r}
levels(x1)
```
**c.**
```{r}
table(x1)[which.min(x1)]
```
**d.**
```{r}
table(x1)[which.min(x1)]/length(x1)
```
**e.**
```{r}
as.numeric(levels(x2)[x2])
x2
```
**f.**
```{r}
typeof(x2)
typeof(x2)
```
## Question 3
**a.**
```{r}
library(dslabs)
```
**b.**
```{r}
glimpse(outlier_example)
```
**c.**
```{r}
m1<- matrix(outlier_example,nrow=50,byrow=TRUE)
```
**d.**
```{r}
ncol(m1)
```
**e.**
```{r}
m1[which.max(m1)]
which.max(m1)
```
## Question 4
**a.**
```{r}
glimpse(greenhouse_gases)
```
**b.**
```{r}
col_f <- factor(greenhouse_gases[,2])
```
**c.**
```{r}
glimpse(greenhouse_gases[,-2])
```
**d.**
```{r}
i_ <- sample((nrow(greenhouse_gases)),10)
```
**e.**
```{r}
greenhouse_gases$concentration[i_] <- NA
sum(is.na(greenhouse_gases["concentration"]))
```
**f.**
```{r}
tf_ <- greenhouse_gases["concentration"]>1000
table(tf_, levels=col_f)
```
## Bonus Question
```{r}
vn_ <- c(-1.1,-0.2, 0, 0.19, 0.89, 1.2)
vc_ <- c("a", "g", "q")
if (vn_ > 0.5 || all(letters %in% vc_)) {
  cat("You fixed it")
}
vn_ > 0.5
all(letters %in% vc_)
```
The following if statement does not work because both the first half and second half of the "or conditional" (represented by ||) is false. The first half of the "or conditional" is asking if every entry in the numeric vector "vn_" is greater than 0.5. The second half of the "or conditional" is asking whether or not every letter of the alphabet is in the character vector "vc_". As neither of these are true, then therefore the "or conditional" will return false and the if statement will not execute. Thus, the if statement will not work.